### 2017-05-12 paper/syntacticSugar.tex
minor

\subsection{Syntactic Sugar} \label{syntacticSugar}
For basic arrows, we have the \code{(***)} combinator (Fig.~\ref{fig:***Img},~\ref{fig:***}) which allows us to combine two arrows \code{arr a b} and \code{arr c d} into an arrow \code{arr (a, c) (b, d)} which does both computations at once. This can easily be translated into a parallel version \code{(|***|)} (Fig.~\ref{fig:|***|}) with the use of \code{parEval2}, but for this we require a backend which has an implementation that does not require any configuration (hence the \code{()} as the conf parameter in Fig.~\ref{fig:|***|}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
(|***|) :: (ArrowChoice arr, ArrowParallel arr (Either a c) (Either b d) ())) =>
	arr a b -> arr c d -> arr (a, c) (b, d)
(|***|) = parEval2 ()
\end{lstlisting}
\caption{Definition of (|***|) - the parallel version of (***)}
\label{fig:|***|}
\end{figure}
% With this we can analogously to the serial \code{&&&}
We define the parallel \code{(|&&&|)} (Fig.~\ref{fig:|&&&|}) in a similar manner to its sequential pendant \code{(&&&)} (Fig.~\ref{fig:&&&Img},~\ref{fig:&&&}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
(|&&&|) :: (ArrowChoice arr, ArrowParallel arr (Either a a) (Either b c) ()) =>
	arr a b -> arr a c -> arr a (b, c)
(|&&&|) f g = (arr $ \a -> (a, a)) >>> f |***| g
\end{lstlisting} % $ %% formatting
\caption{Definition of (|\&\&\&|) - the parallel version of (\&\&\&)}
\label{fig:|&&&|}
\end{figure}

>>>>

\subsection{Syntactic Sugar} \label{syntacticSugar}
For basic arrows, we have the \code{(***)} combinator (Fig.~\ref{fig:***Img},~\ref{fig:***}) which allows us to combine two arrows \code{arr a b} and \code{arr c d} into an arrow \code{arr (a, c) (b, d)} which does both computations at once. This can easily be translated into a parallel version \code{(|***|)} (Fig.~\ref{fig:|***|}) with the use of \code{parEval2}, but for this we require a backend which has an implementation that does not require any configuration (hence the \code{()} as the conf parameter in Fig.~\ref{fig:|***|}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
(|***|) :: (ArrowChoice arr, ArrowParallel arr (Either a c) (Either b d) ())) =>
	arr a b -> arr c d -> arr (a, c) (b, d)
(|***|) = parEval2 ()
\end{lstlisting}
\caption{Definition of (|***|) - the parallel version of (***)}
\label{fig:|***|}
\end{figure}
% With this we can analogously to the serial \code{&&&}
We define the parallel \code{(|\&\&\&|)} (Fig.~\ref{fig:|&&&|}) in a similar manner to its sequential pendant \code{(\&\&\&)} (Fig.~\ref{fig:&&&Img},~\ref{fig:&&&}).
\begin{figure}[h]
\begin{lstlisting}[frame=htrbl]
(|&&&|) :: (ArrowChoice arr, ArrowParallel arr (Either a a) (Either b c) ()) =>
	arr a b -> arr a c -> arr a (b, c)
(|&&&|) f g = (arr $ \a -> (a, a)) >>> f |***| g
\end{lstlisting} % $ %% formatting
\caption{Definition of (|\&\&\&|) - the parallel version of (\&\&\&)}
\label{fig:|&&&|}
\end{figure}

### 2017-05-12 paper/topologySkeletons.tex
minor

\section{Topological Skeletons}
\label{sec:topology-skeletons}
Even though many algorithms can be expressed by parallel maps, some problems require more sophisticated skeletons. The Eden library leverages this problem and already comes with more predefined skeletons, among them a \code{pipe}, a \code{ring} and a \code{torus} implementation \cite{Loogen2012, eden_skel_topology}. These seem like reasonable candidates to be ported to our arrow based parallel Haskell to showcase that we can express such skeletons with Parallel Arrows as well.

@@

The parallel pipe skeleton is semantically equivalent to folding over a list \code{[arr a a]} of arrows with \code{>>>}, but does this in parallel, meaning that the arrows do not have to reside on the same thread/machine. We implement this skeleton using the \code{ArrowLoop} typeclass which gives us the \code{loop :: arr (a, b) (c, b) -> arr a c} combinator which allows us to express loop like computations. For example this
\begin{lstlisting}[frame=htrbl]
loop (arr (\(a, b) -> (b, a:b)))
\end{lstlisting}
,which is the same as
\begin{lstlisting}[frame=htrbl]
loop (arr snd &&& arr (uncurry (:)))
\end{lstlisting}
defines an arrow that takes its input \code{a} and converts it into an infinite stream \code{[a]} of it. Using this to our advantage gives us a first draft of a pipe implementation by plugging in the parallel evaluation call \code{parEvalN conf fs} inside the second argument of \code{&&&} and then only picking the first element of the resulting list with \code{arr last}:
\begin{lstlisting}[frame=htrbl]
pipeSimple :: (ArrowLoop arr, ArrowParallel arr a a conf) =>
	conf -> [arr a a] -> arr a a
pipeSimple conf fs =
	loop (arr snd &&&
		(arr (uncurry (:) >>> lazy) >>> parEvalN conf fs)) >>>
	arr last
\end{lstlisting}
where \code{lazy} is defined as:
\begin{lstlisting}[frame=htrbl]
lazy :: (Arrow arr) => arr [a] [a]
lazy = arr (\ ~(x:xs) -> x : lazy xs)
\end{lstlisting}
Note that here the use of \code{lazy} is essential as without it programs using this definition would never halt. We need to enforce that the evaluation of the input \code{[a]} terminates before passing it into \code{parEvalN}.

>>>>

\FloatBarrier
\section{Topological Skeletons}
\label{sec:topology-skeletons}
Even though many algorithms can be expressed by parallel maps, some problems require more sophisticated skeletons. The Eden library leverages this problem and already comes with more predefined skeletons, among them a \code{pipe}, a \code{ring} and a \code{torus} implementation \cite{Loogen2012, eden_skel_topology}. These seem like reasonable candidates to be ported to our arrow based parallel Haskell to showcase that we can express such skeletons with Parallel Arrows as well.

@@

The parallel pipe skeleton is semantically equivalent to folding over a list \code{[arr a a]} of arrows with \code{>>>}, but does this in parallel, meaning that the arrows do not have to reside on the same thread/machine. We implement this skeleton using the \code{ArrowLoop} typeclass which gives us the \code{loop :: arr (a, b) (c, b) -> arr a c} combinator which allows us to express loop like computations. For example this
\begin{lstlisting}[frame=htrbl]
loop (arr (\(a, b) -> (b, a:b)))
\end{lstlisting}
,which is the same as
\begin{lstlisting}[frame=htrbl]
loop (arr snd &&& arr (uncurry (:)))
\end{lstlisting}
defines an arrow that takes its input \code{a} and converts it into an infinite stream \code{[a]} of it. Using this to our advantage gives us a first draft of a pipe implementation by plugging in the parallel evaluation call \code{parEvalN conf fs} inside the second argument of \code{(\&\&\&)} and then only picking the first element of the resulting list with \code{arr last}:
\begin{lstlisting}[frame=htrbl]
pipeSimple :: (ArrowLoop arr, ArrowParallel arr a a conf) =>
	conf -> [arr a a] -> arr a a
pipeSimple conf fs =
	loop (arr snd &&&
		(arr (uncurry (:) >>> lazy) >>> parEvalN conf fs)) >>>
	arr last
\end{lstlisting}
where \code{lazy} is defined as:
\begin{lstlisting}[frame=htrbl]
lazy :: (Arrow arr) => arr [a] [a]
lazy = arr (\ ~(x:xs) -> x : lazy xs)
\end{lstlisting}
Note that here the use of \code{lazy} is essential as without it programs using this definition would never halt. We need to enforce that the evaluation of the input \code{[a]} terminates before passing it into \code{parEvalN}.
